Considering scale in multilayer animal social networks

### 2020-06-29 paper/manuscript.Rmd
methods edits

We examined scale-dependent multilayer networks motivated by behavioural,
landscape, and spatial ecology to and applied this framework to caribou
(*Rangifer tarandus*) on Fogo Island, Newfoundland, Canada. First, we test to
what degree the social scale of relationships, spatial scale of landscapes, and
temporal scale of analysis influences our ability to interpret complex social
systems. Second, we address how variation in the social, spatial, and temporal
scale improves our understanding of the relationship between social association
and seasonal resource availability. We conclude with potential future
developments and applications of scale in multilayer networks.

@@

## Caribou socioecology
Caribou are gregarious ungulates with fission-fusion dynamics [@Lesmerises_2018]
that display temporal and spatial variation in social networks [@Peignier_2019].
In winter, caribou dig holes, termed craters, to access forage beneath the snow
[@Bergerud_1974]. Cratering is presumed to be a costly behaviour, and as a
result, caribou tend to occupy and re-use craters once they are established.
Importantly, craters exist on the landscape at multiple scales: the crater scale
(a single crater), the feeding area scale (multiple craters in close proximity),
and the winter range scale (all craters within an individual’s range)
[@Mayor_2009]. Access to forage is therefore heterogeneous in winter depending
on snow cover and distribution of craters on the landscape.

## Caribou location data
We used GPS location data collected from Fogo Island caribou between 
`r DT[, format(min(datetime), '%B %Y')]` and `r DT[, format(max(datetime), '%B %Y')]`. 
Adult female caribou were immobilized and fitted with global positioning
system (GPS) collars (Lotek Wireless Inc., Newmarket, ON, Canada, GPS4400M
collars, 1,250 g) as described by Schaefer and Mahoney [-@Schaefer_2013]. Collars were programmed to
collect location fixes every 2 hours. Prior to analyses, we subset GPS fixes
to remove all erroneous and outlier GPS fixes following
[@Bjorneraas_2010]. We did not collar all caribou in the herd; however, the
proportion of marked adult females was ~10% of all adult females and ~5% of all
individuals in the herd. We assumed these individuals were randomly distributed
throughout the population.

@@

### Network layer construction
We generated proximity-based social networks using the R package `spatsoc`
[@Robitaille_2019] in `r substr(version$version.string, 1, 15)` [@R_Core_Team]. 
<!--We assumed
individuals were associating if simultaneous GPS fixes were within the spatial
distance threshold of one another.--> Typically for ungulates and other gregarious
mammals, the ‘chain rule’ is used for group assignment [@Croft_2008]. For GPS
data, the chain rule is applied by assigning a group identifier to the union of
buffered GPS fixes at each time step. As such, individuals in a group are within
the social distance threshold of at least one other though not necessarily all
other individuals [@Robitaille_2019]. Group assignment based on the chain rule
has previously been applied to caribou using a temporal threshold of 5 minutes
and a social distance threshold of 50 m [@Peignier_2019; @Lesmerises_2018]. In
all networks, individual caribou were represented as nodes and associations between
individuals in a given network layer were represented as intralayer
edges.

@@

#### Number of observations
Studies of social network analysis vary in the number and frequency of
observations as well as the data collection technique used to generate networks,
ranging from directly observed to remotely sensed [@Davis_2018; @Webber_2019].
Remotely sensed GPS data is commonly collected at a fixed rate, e.g. every $x$
minutes or hours, continuously throughout the study period. Fix rate is a
reflection of number of observations an individual would have been observed and
recorded in traditional ethological studies. To investigate the influence of
number of observations, we generated multilayer networks of social association
between `r DT[, uniqueN(get(idcol))]` individuals across three habitat layers
(foraging, forest, and open) using $n$ randomly selected observations. We first
generated a maximum of 1000 timesteps and iteratively used $n$ of these ($n$ =
10 - 1000 observations by steps of 10), to ensure previously included timesteps,
and resulting associations, were accumulated with subsequent observations to
mimic collection of observational data. Within each multilayer network with $n$
observations, we calculated individual graph strength and multidegree. We
calculated edge overlap across the entire series of networks to determine
proportion of total edges as number of observations increased.

>>>>

We examined scale-dependent multilayer networks motivated by behavioural,
landscape, and spatial ecology to and applied this framework to caribou
(*Rangifer tarandus*). First, we test to
what degree the social scale of relationships, spatial scale of landscapes, and
temporal scale of analysis influences our ability to interpret complex social
systems. Second, we address how variation in the social, spatial, and temporal
scale improves our understanding of the relationship between social association
and seasonal resource availability. We conclude with potential future
developments and applications of scale in multilayer networks.

@@

## Caribou socioecology
Caribou are gregarious ungulates with fission-fusion dynamics [@Lesmerises_2018]
that display temporal and spatial variation in social networks [@Peignier_2019].
In winter, caribou dig holes, termed craters, to access forage beneath the snow
[@Bergerud_1974]. Cratering is presumed to be a costly behaviour, and as a
result, caribou tend to occupy and re-use craters once they are established.
Importantly, craters exist on the landscape at multiple scales: the crater scale
(a single crater), the feeding area scale (multiple craters in close proximity),
and the winter range scale (all craters within an individual’s range)
[@Mayor_2009]. Access to forage is therefore heterogeneous in winter and depends
on snow cover and distribution of craters on the landscape.

## Caribou location data
We used GPS location data collected from caribou on Fogo Island, Newfoundland, Canada between 
`r DT[, format(min(datetime), '%B %Y')]` and `r DT[, format(max(datetime), '%B %Y')]`. 
Adult female caribou were immobilized and fitted with global positioning
system (GPS) collars (Lotek Wireless Inc., Newmarket, ON, Canada, GPS4400M
collars, 1,250 g) as described by Schaefer and Mahoney [-@Schaefer_2013]. Collars were programmed to
collect location fixes every 2 hours. Prior to analyses, we subset GPS fixes
to remove all erroneous and outlier GPS fixes following
[@Bjorneraas_2010]. We did not collar all caribou in the herd; however, the
proportion of marked adult females was ~10% of all adult females and ~5% of all
individuals in the herd. We assumed these individuals were randomly distributed
throughout the population.

@@

### Network layer construction
We generated proximity-based social networks using the R package `spatsoc`
[@Robitaille_2019] in `r substr(version$version.string, 1, 15)` [@R_Core_Team]. 
Typically for ungulates and other gregarious
mammals, the ‘chain rule’ is used for group assignment [@Croft_2008]. For GPS
data, the chain rule is applied by assigning a group identifier to the union of
buffered GPS fixes at each time step. As such, individuals in a group are within
the social distance threshold of at least one other though not necessarily all
other individuals [@Robitaille_2019]. Group assignment based on the chain rule
has previously been applied to caribou using a temporal threshold of 5 minutes
and a social distance threshold of 50 m [@Peignier_2019; @Lesmerises_2018]. In
all networks, individual caribou were represented as nodes and associations between
individuals in a given network layer were represented as intralayer
edges.

@@

#### Number of observations
Studies of social network analysis vary in the number and frequency of
observations as well as the data collection technique used to generate networks,
ranging from directly observed to remotely sensed [@Davis_2018; @Webber_2019].
Remotely sensed GPS data are commonly collected at a fixed rate, e.g. every $x$
minutes or hours, continuously throughout the study period. Fix rate is a
reflection of number of observations an individual would have been observed and
recorded in traditional ethological studies. To investigate the influence of
number of observations, we generated multilayer networks of social association
between `r DT[, uniqueN(get(idcol))]` individuals across three habitat layers
(foraging, forest, and open) using $n$ randomly selected observations. We first
generated a maximum of 1000 timesteps and iteratively used $n$ of these ($n$ =
10 - 1000 observations by steps of 10), to ensure previously included timesteps,
and resulting associations, were accumulated with subsequent observations to
mimic collection of observational data. Within each multilayer network with $n$
observations, we calculated individual graph strength and multidegree. We
calculated edge overlap across the entire series of networks to determine
proportion of total edges as number of observations increased.
